{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-21 thesis/landslide.tex
final pre-send-to-committee revisions to education chapater
i don't know if blacklist/whitelist is truly racist but it cost literally 0 dollars and 0 cents to change ur language just to be safe

\begin{itemize}
	\item {\tt within\_function FUNC} - adds {\tt FUNC} to a whitelist of functions required to appear in the current stack trace before identifying a preemption point (see \sect{\ref{sec:landslide-pps}})
	\item {\tt without\_function FUNC} - as above, but a blacklist instead of a whitelist
	\item {\tt within\_user\_function FUNC} - as two above but finds the function in the userspace test program rather than the kernel code.
	\item {\tt without\_user\_function FUNC} - difference to two above same as stated one above.
	\item {\tt data\_race ADDR TID LAST\_CALL CURRENT\_SYSCALL} - specifies a data-race preemption point.
		\begin{itemize}
			\item {\tt ADDR} shall be the code address (in hex) of the racing address,
			{\em before} the execution of which a preemption will be issued.
			\item {\tt TID} indicates a thread ID required to be running for this data race.
				To specify data race PPs across all threads at once, set {\tt FILTER\_DRS\_BY\_TID=0} (see next section).
			\item {\tt LAST\_CALL} indicates a code address required to be the site of the last {\tt call} instruction executed
				(similar to specifying a stack trace, but using a full stack trace here degrades performance too much),
				or 0 to not use this feature.
				From personal experience I found this option rather useless and recommend always supplying 0.
				% TODO: fix this section ref
				For further discussion see \sect{\ref{sec:quicksand-pps}}.
			\item {\tt CURRENT\_SYSCALL} indicates the system call number if a user-space data race comes from within a kernel system call which accesses user memory (Pebbles only).
				Usually 0 (i.e., not in kernel code) but {\tt deschedule}'s system call number is common as well.
		\end{itemize}
	\item {\tt input\_pipe FILENAME} - FIFO file used for receiving messages from Quicksand (e.g. to suspend or resume execution).
		Requires {\tt id\_magic} option to be set (next section below).
		The odds that a human user will find spiritual enlightenment through using this option by hand are infinitesimal.
	\item {\tt output\_pipe FILENAME} - as above but for sending messages.
\end{itemize}

@@

\subsubsection{Heap checking}

@@

{\tt check\_withins()} implements the whitelist/blacklist behaviour for the former two of those commands
(in a similar manner to the Preemption Sealing technique described in prior work \cite{sealing}).
It invokes the stack tracer (\sect{\ref{sec:landslide-stacktrace}})
for a list of which functions are on the call stack
(hence the importance of the stack tracer's complex logic to not miss any frames
even when interrupts or system calls are involved).
Then, to determine if the current program state should be considered a valid preemption point,
or whether it should be rejected,
it compares each {\tt within} or {\tt without\_function} directive in the following sequence-dependent manner:

\begin{itemize}
	\item If no {\tt within\_function} commands are given, operate in ``blacklist mode'':
		the preemption point is by default valid as long as no {\tt without\_function} calls reject it.
		Otherwise, operate in ``whitelist mode'':
		the preemption point is by default rejected unless at least one {\tt within\_function} directive matches.
	\item Subject to the above, find the sequentially-last {\tt *\_function} directive
		(static preemption point commands considered before dynamic ones)
		which matches any function in the stack trace.
		If {\tt within}, accept the preemption point; if {\tt without}, reject it.
\end{itemize}

@@

		No {\tt without}s present, so behaves as a whitelist,
		rejecting stack trace 1 (not in {\tt thr\_join()}),
		accepting stack trace 2 (in {\tt thr\_join()}).
	\item
		{\tt without\_user\_function cond\_wait}

		No {\tt within}s present, so behaves as a blacklist,
		accepting stack trace 1 (not in {\tt cond\_\allowbreak{}wait()}),
		rejecting stack trace 2 (in {\tt cond\_wait()}).
	\item
		{\tt without\_user\_function main},
		{\tt within\_user\_function mutex\_lock}

>>>>

\begin{itemize}
	\item {\tt within\_function FUNC} - adds {\tt FUNC} to an allowlist of functions required to appear in the current stack trace before identifying a preemption point (see \sect{\ref{sec:landslide-pps}})
	\item {\tt without\_function FUNC} - as above, but a denylist instead of an allowlist
	\item {\tt within\_user\_function FUNC} - as two above but finds the function in the userspace test program rather than the kernel code.
	\item {\tt without\_user\_function FUNC} - difference to two above same as stated one above.
	\item {\tt data\_race ADDR TID LAST\_CALL CURRENT\_SYSCALL} - specifies a data-race preemption point.
		\begin{itemize}
			\item {\tt ADDR} shall be the code address (in hex) of the racing address,
			{\em before} the execution of which a preemption will be issued.
			\item {\tt TID} indicates a thread ID required to be running for this data race.
				To specify data race PPs across all threads at once, set {\tt FILTER\_DRS\_BY\_TID=0} (see next section).
			\item {\tt LAST\_CALL} indicates a code address required to be the site of the last {\tt call} instruction executed
				(similar to specifying a stack trace, but using a full stack trace here degrades performance too much),
				or 0 to not use this feature.
				From personal experience I found this option rather useless and recommend always supplying 0.
				% TODO: fix this section ref
				For further discussion see \sect{\ref{sec:quicksand-pps}}.
			\item {\tt CURRENT\_SYSCALL} indicates the system call number if a user-space data race comes from within a kernel system call which accesses user memory (Pebbles only).
				Usually 0 (i.e., not in kernel code) but {\tt deschedule}'s system call number is common as well.
		\end{itemize}
	\item {\tt input\_pipe FILENAME} - FIFO file used for receiving messages from Quicksand (e.g. to suspend or resume execution).
		Requires {\tt id\_magic} option to be set (next section below).
		The odds that a human user will find spiritual enlightenment through using this option by hand are infinitesimal.
	\item {\tt output\_pipe FILENAME} - as above but for sending messages.
\end{itemize}

@@

\subsubsection{Heap checking}
\label{sec:landslide-valgrind-mode}

@@

{\tt check\_withins()} implements the allow/denylist behaviour for the former two of those commands
(in a similar manner to the Preemption Sealing technique described in prior work \cite{sealing}).
It invokes the stack tracer (\sect{\ref{sec:landslide-stacktrace}})
for a list of which functions are on the call stack
(hence the importance of the stack tracer's complex logic to not miss any frames
even when interrupts or system calls are involved).
Then, to determine if the current program state should be considered a valid preemption point,
or whether it should be rejected,
it compares each {\tt within} or {\tt without\_function} directive in the following sequence-dependent manner:

\begin{itemize}
	\item If no {\tt within\_function} commands are given, operate in ``denylist mode'':
		the preemption point is by default valid as long as no {\tt without\_function} calls reject it.
		Otherwise, operate in ``allowlist mode'':
		the preemption point is by default rejected unless at least one {\tt within\_function} directive matches.
	\item Subject to the above, find the sequentially-last {\tt *\_function} directive
		(static preemption point commands considered before dynamic ones)
		which matches any function in the stack trace.
		If {\tt within}, accept the preemption point; if {\tt without}, reject it.
\end{itemize}

@@

		No {\tt without}s present, so behaves as an allowlist,
		rejecting stack trace 1 (not in {\tt thr\_join()}),
		accepting stack trace 2 (in {\tt thr\_join()}).
	\item
		{\tt without\_user\_function cond\_wait}

		No {\tt within}s present, so behaves as a denylist,
		accepting stack trace 1 (not in {\tt cond\_\allowbreak{}wait()}),
		rejecting stack trace 2 (in {\tt cond\_wait()}).
	\item
		{\tt without\_user\_function main},
		{\tt within\_user\_function mutex\_lock}
